---
title: "BiocBaseUtils Quick Start"
author: "Bioconductor Core Team"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  BiocStyle::html_document:
    number_sections: yes
    toc: true
vignette: >
  %\VignetteIndexEntry{BiocBaseUtils Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# BiocBaseUtils

The `BiocBaseUtils` package provides a suite of helper functions designed to
help developers. Currently, it covers three topics often encountered during
the development process.

1. Assertions - Type checks for logical, character, and numeric inputs
2. Slot replacement - Replacing the value of object slots
3. `show` method - Limiting the output of internal components of a class

# Installation

Install the package directly from Bioconductor:

```{r,eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("BiocBaseUtils")
```

# Load Package

```{r,include=TRUE,results="hide",message=FALSE,warning=FALSE}
library(BiocBaseUtils)
```

# Assertions

We provide a number of functions that helps the developer establish the
type of class of a particular object. These include `integer`, `numeric`,
`character`, and `logical`; types often used in R / Bioconductor.

## Logical

```{r}
isTRUEorFALSE(TRUE)
isTRUEorFALSE(FALSE)
isTRUEorFALSE(NA, na.ok = TRUE)
```

## Character

```{r}
isScalarCharacter(LETTERS)
isScalarCharacter("L")
isCharacter(LETTERS)
isCharacter(NA_character_, na.ok = TRUE)
isZeroOneCharacter("")
isZeroOneCharacter("", zchar = TRUE)
```

## Numeric

```{r}
isScalarInteger(1L)
isScalarInteger(1)

isScalarNumber(1)
isScalarNumber(1:2)
```

# Slot replacement

This function is often used in packages that establish formal S4 classes.
When updating the value of a slot, one often uses the `setSlots` function.

```{r}
setClass("A", representation = representation(slot1 = "numeric"))
aclass <- new("A", slot1 = 1:10)
aclass
```

Now we use the `setSlots` function to update the values in the object.

```{r}
aclass <- setSlots(aclass, slot1 = 11:20)
aclass
```

Note that `setSlots` provides the same functionality as
`BiocGenerics:::replaceSlots` but is more consistent with Bioconductor the
setter and getter language.

# `show` method

The `selectSome` function allows the developer to display a limited amount of
information from a developed class. Note that the use of the `@` here is due
to the minimal implementation in the examples provided. The developer should
always provide an interface to access the internal components of the class
via an 'accessor' function.

```{r}
setMethod("show", signature = "A", function(object) {
    s1info <- getElement(object, "slot1")
    cat("A sequence:", selectSome(s1info))
})
aclass
```

# Contributing

`BiocBaseUtils` is a work in progress and we welcome contributions. There
are quite a few often-used utility functions that are yet to be included in the
package. We would like to keep the dependencies in this package minimal;
therefore, contributions should mostly use base R.

# Session Info

```{r}
sessionInfo()
```

Please report minimally reproducible bugs at our [github issue page][].

[github issue page]: https://github.com/Bioconductor/BiocBaseUtils/issues
